---
title: "Correlated numerical"
format:
  html:
    theme: sketchy
    fontsize: 1.1em
    linestretch: 1.5
    toc: true
    toc-location: left-body
    toc-depth: 2
    number-sections: true
    number-depth: 3
---

Report prepared by: 
1. Aina Syazwani bt Aziz 
2. Nur Shuhaiza bt Supian
3. Muhammad Aminuddin Bin Che Hashim

# Introduction

In healthcare settings, particularly in hospital environments, nurses' well-being and performance are closely linked to their sleep quality. Sleep disturbances among nurses have been associated with reduced job performance, increased risk of errors, and compromised patient safety. To better understand the factors influencing sleep quality, it is important to consider the organizational structure within which nurses operate. Specifically, nurses are nested within clinical departments, which are themselves nested within larger institutional frameworks such as hospitals. This hierarchical structure creates dependencies among observations that violate the assumptions of traditional single-level regression models.

In this study, multilevel modeling is employed to analyze sleep scores among nurses working in two tertiary hospitals: Hospital Universiti Sains Malaysia (HUSM) and Hospital Raja Perempuan Zainab II (HRPZ II). The data exhibit a three-level structure: individual nurses (Level 1) are nested within departments (Level 2), which are in turn nested within hospitals (Level 3). A multilevel modeling approach allows for the partitioning of variance in sleep scores across these levels, enabling a more accurate understanding of both individual and contextual factors that may influence sleep quality.

The primary outcome variable, sleep score, is continuous and reflects the overall sleep quality of each nurse. By incorporating random intercepts at both the department and hospital levels, the model accounts for unobserved heterogeneity that may arise from departmental culture, workload, shift patterns, and institutional policies. Fixed effects can be included to assess how individual characteristics (e.g., age, shift type, workload) and group-level factors (e.g., department type or hospital) impact sleep quality.

Multilevel analysis not only improves the precision of estimates by accounting for clustering but also offers insights into how much of the variability in sleep outcomes is attributable to departments or hospitals. This approach supports more targeted and level-specific interventions to enhance nurse well-being and, by extension, healthcare quality.

# Methodologies
## About the Data
The data comprises 462 participants distributed across two tertiary hospitals in Kelantan. The data was collected for MPH study. 

The dataset is structured hierarchically, with individual nurses (Level 1) nested within hospital departments (Level 2), which are further nested within tertiary hospitals (Level 3). Nurses were drawn from multiple departments, including emergency, pediatrics, surgical, and medical wards—each with unique work demands and environmental factors potentially influencing sleep.

The outcome of interest is a continuous sleep score (total_score), which quantitatively measures sleep quality. Explanatory variables include both individual-level characteristics (e.g., age, gender, marital status, position, years of service, work hours, presence of comorbidities) and department- or hospital-level factors (e.g., type of department, institutional policies, shift schedules).

# Variables
The data consist of several variables. The following are list of variables in the data:

a) id: participants identification
b) hospital: tertiary hospital (Hospital Universiti Sains Malaysia or Hospital Raja Perempuan Zainab II)
c) age: participants' age in years (continuous variable)
d) gender: participants gender (perempuan or lelaki)
e) race: participant race (Melayu, Cina etc)
f) status: bujang, berkahwin, bercerai (samaada hidup atau mati)
g) child: number of participants' child (continuous variable)
h) education: Degree/Master/PHD, Diploma dengan Post-Basic, Diploma/Kolej, Sijil (SPM/STPM)
i) position: Jururawat Masyarakat, Jururawat, Penyelia Jururawat, Ketua Jururawat
j) department: Emergency, Pediatrics, Surgical or Medical
k) schedule: shift or non-shift
l) total_service: total years of services (continuos variable)
m) work_hours: hours of working (continuos variable)
n) commuting_injury: history of commuting injury (yes/no)
o) medication_error: history of medication error (yes/no)
p) comorbid: yes/no
q) total_score: total score from validated questionnaire (continuos variable)
r) sleep_quality: poor or good


# Load packages

### General packages
```{r}
library(tidyverse)
library(gtsummary)
library(kableExtra)
library(plotly)
library(haven)
library(broom.mixed)
library(DT)
library(readxl)
```

### Packages for multilevel
```{r}
library(lme4)
library(lmerTest)
```

# Read data
```{r}
sleep <- read_excel("sleep2.xlsx")
glimpse(sleep)
```

# Data wrangling
```{r}
sleep<-sleep %>% mutate_if(is.character,~ as_factor(.))
glimpse(sleep)
```
# Exploratory Data Analysis
```{r}
summary(sleep)
sleep %>% select(-id) %>% 
  tbl_summary(
    statistic = list(all_continuous() ~
                                 "{mean} ({sd})")) %>% 
  bold_labels()

```

Plot
```{r}
sleep %>% ggplot(aes(x=hospital, y = total_score)) +
  geom_point() +
  geom_smooth(method = lm)
```

```{r}
sleep %>% ggplot(aes(x=hospital, y= total_score,
                     col= department, group= department)) +
  geom_point() +
  geom_smooth(method = lm)
```

# Single level analysis

Single level analysis is used to model a single-level analysis (i.e., no clustering or grouping is accounted for) of the numerical outcome (total_score). 

```{r}
sleep.lm<- lm(total_score ~ 1, data = sleep)
summary(sleep.lm)
```
```{r}
tidy(sleep.lm)
```

# Multilevel analysis
We will use lme4 packages and start with constant model or also known as null model. There will be no explanatory variables. We set the estimation using maximum likelihood estimates(MLE).

The null model will be name as m0. the random effect is due to hospital. 

This is a random intercept with constant-only model.

```{r}
m0<- lmer(total_score ~ 1 + (1|hospital), data = sleep, REML= FALSE)
summary(m0)
```
The overall mean score (across hospitals) is estimated as 7.15. The mean score for hospital j is estimated as 7.15 + U0j where U0j is the hospital residual.

```{r}
tidy(m0) %>% 
  kbl() %>% 
  kable_styling()
```
Adding another level (department) into the model.
```{r}
m1<- lmer(total_score ~ 1 + (1|hospital)+ (1|hospital:department),
          data = sleep, REML = FALSE)
summary(m1)
```
This model consist of random intercepts for each hospital and random intercepts for departments nested within hospital.

# Random intercept model

## Explanatory variable

## Prediction

## Plot

## Variance

### Between hospital

### Within hospital

# Random slope model

## Model

## Fitted value

# Comparing model (RI vs RS)

# Interpretation
