---
title: "Multinomial Outcomes"
format:
  html:
    theme: sketchy
    fontsize: 1.1em
    linestretch: 1.5
    toc: true
    toc-location: left-body
    toc-depth: 2
    number-sections: true
    number-depth: 3
---

Report prepared by: 
1. Aina Syazwani bt Aziz 
2. Nur Shuhaiza bt Supian
3. Muhammad Aminuddin Bin Che Hashim

# Introduction



# Load packages
## General packages
```{r}
library(tidyverse)
library(broom)
library(gtsummary)
library(broom)
library(kableExtra)
library(haven)
```

# About the dataset
The dataset consist of the following variables;

# Load the dataset
```{r}
data<- read_csv("datamssm_a.csv")
glimpse(data)
```
# Data wrangling
Change character variable to factor
```{r}
data<- data %>% mutate_if(is.character, ~as_factor(.))
glimpse(data)
```
# Exploratory Data Analysis
```{r}
summary(data)
```
There are several data missing in the variable. Hence, we need to handle these missing data prior to continuing with analysis.

## Handling missing data
In this practice, we remove any row with missing data. 
```{r}
clean_data <- na.omit(data)
summary(clean_data)
```
```{r}
glimpse(clean_data)
clean_data %>% select(-codesub) %>% 
  tbl_summary(by= dmdx,
              statistic = list(all_continuous()~ "{mean} ({sd})"))
```
## Categorize fbs into 3 levels
```{r}
# Categorize FBS based on Malaysian clinical cut-offs
clean_data <- clean_data %>%
  mutate(cat_fbs = case_when(
    fbs < 5.6 ~ "Normal",
    fbs >= 5.6 & fbs <= 6.9 ~ "Prediabetes",
    TRUE ~ "Diabetes")) %>%
  mutate(cat_fbs = factor(cat_fbs, levels = c("Normal", "Prediabetes", "Diabetes")))

# Check result
table(clean_data$cat_fbs)

```
## Checking level of cat_fbs
```{r}
levels(clean_data$cat_fbs)
```
## Re-level for cat_fbs
We will use VGAM package for further multinomial analysis. As the packages compare biggest value to the lowest, we need to rearrage the level of cat_fbs. We will make "Normal" as the reference category.
```{r}
clean_data<- clean_data %>% 
  mutate(cat_fbs2 = fct_relevel(cat_fbs,
                                c("Diabetes", "Prediabetes", "Normal")))

# compare the data
summary(clean_data$cat_fbs); summary(clean_data$cat_fbs2)
```
```{r}
# recheck releveling
levels(clean_data$cat_fbs2)
```
# Multinomial analysis
We will be using VGAM packages to run multinomial logistic regression. The model that we want to estimate for are:
1. group 1 vs group 3: Diabetes vs Normal
2. group 2 vs group 3: Prediabetes vs Normal

The reference group here, is Normal group.

## Load the packages
```{r}
library(VGAM)
```
## Model with univariable
Gender
```{r}
mod1<- vglm(cat_fbs2~ gender, multinomial, data = clean_data)
summary(mod1)
```
Intercept 1: for outcome cat_fbs2 = diabetes vs normal
Intercept 2: for outcome cat_fbs2 = prediabetes vs normal
gendermale:1 for outcome cat_fbs 2 = diabetes vs normal
gendermale:2 for outcome cat_fbs2 = prediabetes vs normal

Age
```{r}
mod2<- vglm(cat_fbs2 ~ age, multinomial, data = clean_data)
summary(mod2)
```
interpretation
