---
title: "Multinomial Outcomes"
format:
  html:
    theme: sketchy
    fontsize: 1.1em
    linestretch: 1.5
    toc: true
    toc-location: left-body
    toc-depth: 2
    number-sections: true
    number-depth: 3
---

Report prepared by: 
1. Aina Syazwani bt Aziz 
2. Nur Shuhaiza bt Supian
3. Muhammad Aminuddin Bin Che Hashim

# Introduction
Multinomial logistic regression is a statistical method used to model relationships between a polytomous (i.e., having more than two categories) nominal outcome variable and one or more independent variables. Unlike binary logistic regression, which is limited to dichotomous outcomes, multinomial logistic regression accommodates outcome variables with three or more unordered categories. This method estimates the probability of each category relative to a reference group by fitting a series of binary logistic regressions, producing a set of coefficients that reflect the log-odds of membership in each non-reference category. It is particularly useful in public health and epidemiological research where outcomes such as disease classification, behavioral responses, or clinical stages are categorized without a natural ordering. The model can incorporate both continuous and categorical predictors, enabling comprehensive analysis of complex associations within population health datasets.

# Load packages
## General packages
```{r}
library(tidyverse)
library(broom)
library(gtsummary)
library(broom)
library(kableExtra)
library(haven)
```

# About the dataset
This dataset derived from a health or metabolic syndrome screening study. It includes anthropometric, biochemical, and lifestyle variables. The primary outcome of interest for your analysis is fasting blood sugar (fbs).

The dataset consist of the following variables;
codesub: Participant code or ID.
age: Age in years (numeric).
hpt: Hypertension status — “yes” or “no”.
smoking: Smoking status — e.g., “never smoked”, “still smoking”.
dmdx: Diabetes diagnosis — “yes” or “no”.
height: Height in meters (numeric).
weight: Weight in kilograms (numeric).
waist: Waist circumference in centimeters (numeric).
hip: Hip circumference in centimeters (numeric).
msbpr: Measured systolic blood pressure (numeric).
mdbpr: Measured diastolic blood pressure (numeric).
hba1c: Hemoglobin A1c (%) (numeric).
fbs: Fasting blood sugar in mmol/L — this is the outcome variable.
mogtt1h: Blood glucose level at 1 hour during OGTT (Oral Glucose Tolerance Test) (mmol/L).
mogtt2h: Blood glucose level at 2 hours during OGTT (mmol/L).
totchol: Total cholesterol (mmol/L).
ftrigliz: Fasting triglycerides (mmol/L).
hdl: High-density lipoprotein cholesterol (mmol/L).
ldl: Low-density lipoprotein cholesterol (mmol/L).
gender: Gender — “male” or “female”.
crural: Locality — “rural” or possibly “urban”.

# Load the dataset
```{r}
data<- read_csv("datamssm_a.csv")
glimpse(data)
```
# Data wrangling
Change character variable to factor
```{r}
data<- data %>% mutate_if(is.character, ~as_factor(.))
glimpse(data)
```
# Exploratory Data Analysis
```{r}
summary(data)
```
There are several data missing in the variable. Hence, we need to handle these missing data prior to continuing with analysis.

## Handling missing data
In this practice, we remove any row with missing data. 
```{r}
clean_data <- na.omit(data)
summary(clean_data)
```
```{r}
glimpse(clean_data)
clean_data %>% select(-codesub) %>% 
  tbl_summary(by= dmdx,
              statistic = list(all_continuous()~ "{mean} ({sd})"))
```
## Categorize fbs into 3 levels
```{r}
# Categorize FBS based on Malaysian clinical cut-offs
clean_data <- clean_data %>%
  mutate(cat_fbs = case_when(
    fbs < 5.6 ~ "Normal",
    fbs >= 5.6 & fbs <= 6.9 ~ "Prediabetes",
    TRUE ~ "Diabetes")) %>%
  mutate(cat_fbs = factor(cat_fbs, levels = c("Normal", "Prediabetes", "Diabetes")))

# Check result
table(clean_data$cat_fbs)

```
## Checking level of cat_fbs
```{r}
levels(clean_data$cat_fbs)
```
## Re-level for cat_fbs
We will use VGAM package for further multinomial analysis. As the packages compare biggest value to the lowest, we need to rearrage the level of cat_fbs. We will make "Normal" as the reference category.
```{r}
clean_data<- clean_data %>% 
  mutate(cat_fbs2 = fct_relevel(cat_fbs,
                                c("Diabetes", "Prediabetes", "Normal")))

# compare the data
summary(clean_data$cat_fbs); summary(clean_data$cat_fbs2)
```
```{r}
# recheck releveling
levels(clean_data$cat_fbs2)
```
# Multinomial analysis
We will be using VGAM packages to run multinomial logistic regression. The model that we want to estimate for are:
1. group 1 vs group 3: Diabetes vs Normal
2. group 2 vs group 3: Prediabetes vs Normal

The reference group here, is Normal group.

## Load the packages
```{r}
library(VGAM)
```
## Model with univariable
###Gender
```{r}
mod1<- vglm(cat_fbs2~ gender, multinomial, data = clean_data)
summary(mod1)
```
Intercept 1: for outcome cat_fbs2 = diabetes vs normal
(Intercept):1 = -2.13590
This value represents the log-odds of being in the Diabetes group rather than the Normal group for individuals in the reference category, which is female.
The coefficient is statistically significant (p < 0.001), indicating a meaningful difference.
Since the log-odds value is negative, it suggests that females have a low likelihood of being diabetic compared to having normal fasting blood sugar levels.

Intercept 2: for outcome cat_fbs2 = prediabetes vs normal
(Intercept):2 = -1.29294
This coefficient represents the log-odds of being in the Prediabetes group rather than the Normal group for individuals in the reference category, which is female.
It is statistically significant (p < 0.001), indicating a strong association.
The negative value suggests that females have lower odds of being prediabetic compared to having normal fasting blood sugar levels.

gendermale:1 for outcome cat_fbs 2 = diabetes vs normal
gendermale:1 = 0.07934 (p = 0.53545)
This coefficient reflects the change in log-odds of being in the Diabetes group versus the Normal group for males compared to females.
The result is not statistically significant (p > 0.05), indicating that gender does not have a meaningful association with the likelihood of diabetes compared to normal fasting blood sugar levels in this model.

gendermale:2 for outcome cat_fbs2 = prediabetes vs normal
This coefficient represents the change in log-odds of being in the Prediabetes group compared to the Normal group for males relative to females.
The association is statistically significant, indicating that males are significantly more likely to be prediabetic than females when compared to having normal fasting blood sugar levels.

```{r}
exp(0.23841)
```
So, males have 1.27 times the odds of being prediabetic vs normal compared to females.

###Age
```{r}
mod2<- vglm(cat_fbs2 ~ age, multinomial, data = clean_data)
summary(mod2)
```

age:1 for outcome cat_fbs 2 = diabetes vs normal 
age:1          0.030088
Each 1-year increase in age is associated with a 3.1% increase in odds of being diabetic vs normal

age:2 for outcome cat_fbs2 = prediabetes vs normal
age:2          0.030949
Each 1-year increase in age is associated with a 3.2% increase in odds of being prediabetic vs normal

Age is significantly associated with both diabetes and prediabetes, compared to normal FBS. 
For each additional year of age:
The odds of having diabetes (vs normal FBS) increase by ~3.1%
The odds of having prediabetes (vs normal FBS) increase by ~3.2%

This indicates that older individuals are more likely to have impaired fasting blood sugar.

##Rural vs urban
```{r}
mod3<- vglm(cat_fbs2 ~ crural, multinomial, data = clean_data)
summary(mod3)
```
## Inferences


