---
title: "Survival semiparametric"
format:
  html:
    theme: sketchy
    fontsize: 1.1em
    linestretch: 1.5
    toc: true
    toc-location: left-body
    toc-depth: 2
    number-sections: true
    number-depth: 3
---

Report prepared by: 1. Aina Syazwani bt Aziz 2. Nur Shuhaiza bt Supian 3. Muhammad Aminuddin Bin Che Hashim ---

# Introduction

# Load packages

## General packages

```{r}
library(tidyverse)
library(broom)
library(gtsummary)
library(broom)
library(psych)
library(corrplot)
library(tidyr)
library(survival)
library(survminer)
```

# About the dataset

The dataset consist of the following variables;

# Load the dataset

```{r}
library(haven)
stroke_fatality <- read_dta("stroke_fatality.dta")
View(stroke_fatality)
summary(stroke_fatality)
```

# Data wrangling

Change character variable to factor

```{r}
stroke_fatality <- stroke_fatality %>% mutate_if(is.labelled, ~as_factor(.))
glimpse(stroke_fatality)

```

# Exploratory Data Analysis

```{r}
summary(stroke_fatality)
```

There are several data missing in the variable. Hence, we need to handle these missing data prior to continuing with analysis.

## Handling missing data

In this practice, we remove any row with missing data.

```{r}
clean_data <- na.omit(data)
summary(clean_data)
```

```{r}
glimpse(clean_data)
clean_data %>% select(-codesub) %>% 
  tbl_summary(by= dmdx,
              statistic = list(all_continuous()~ "{mean} ({sd})"))
```
